The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor Couette flow

### 2016-09-30 paper/thingap_mri_paper.tex
latex corrections

The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to $k$ \citep{Hoyle:2006}. In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane.  In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well. Figure~\ref{fig:u_saturation} shows the saturated mean radial profiles of $u_0 - u_{\phi} = -q \Omega_0 x - u_{\phi}$ and each term in the steady state force balance (i.e. equation~(\ref{eqset2} with $\d_t u_{1y} = 0$). In the bulk of the fluid away from the boundary layers, the saturated state shows reduced shear, with little diffusive contribution. However, the steep scaling of transport with $\Pm$

Figure~\ref{fig:b_saturation} shows $B_z$ and the terms corresponding to steady state inductive balance (equation~(\ref{eqset2} with $\d_t B_{y} = 0$). Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers. Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition. \citet{Ebrahimi:2009} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean. They considered two important cases, one in which the mean flow was forced to remain at its initial, quasi-Keplerian state for all time, and one in which the background flow was allowed to evolve. This is a crucial difference between the shearing box and our narrow-gap TC flow: perturbations in our simulation can adjust the background flow, whereas in a shearing box, the shear periodicity forbids perturbations from affecting the mean flow. In the case with a freely evolving background flow, \citet{Ebrahimi:2009} found that a saturated state quite similar to ours: field pushed to the boundaries, and a reduction in shear in the bulk of the flow. Their flows have less pronounced boundary layers, likely because of their much larger $\Pm = 0.1 -1$.

>>>>

The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to $k$ \citep{Hoyle:2006}. In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane.  In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well. Figure~\ref{fig:u_saturation} shows the saturated mean radial profiles of $u_0 - u_{\phi} = -q \Omega_0 x - u_{\phi}$ and each term in the steady state force balance (i.e. equation~(\ref{eqset2} with $\partial_t u_{1y} = 0$). In the bulk of the fluid away from the boundary layers, the saturated state shows reduced shear, with little diffusive contribution. However, the steep scaling of transport with $\Pm$

Figure~\ref{fig:b_saturation} shows $B_z$ and the terms corresponding to steady state inductive balance (equation~(\ref{eqset2} with $\partial_t B_{y} = 0$). Here, the instability acts to push magnetic field toward the boundaries in both the bulk and the boundary layers. Note also that the total magnetic energy in the $B_z$ component is significantly larger than the initial condition. \citet{Ebrahimi:2009} considered the saturation of a single, strongly super-critical MRI mode allowed to interact non-linearly only with itself and the mean. They considered two important cases, one in which the mean flow was forced to remain at its initial, quasi-Keplerian state for all time, and one in which the background flow was allowed to evolve. This is a crucial difference between the shearing box and our narrow-gap TC flow: perturbations in our simulation can adjust the background flow, whereas in a shearing box, the shear periodicity forbids perturbations from affecting the mean flow. In the case with a freely evolving background flow, \citet{Ebrahimi:2009} found that a saturated state quite similar to ours: field pushed to the boundaries, and a reduction in shear in the bulk of the flow. Their flows have less pronounced boundary layers, likely because of their much larger $\Pm = 0.1 -1$.
